Introduction to\\ Swedish Civil Contingencies Agency (MSB) methodological support for introducing \ac{isms}.

### 2019-03-20 msbframework/msbintro/msbintro.tex
Fixed some formatting

\DeclareAcronym{isms}
{
  short = ISMS,
  long = Information Security Management Systems,
}

@@

\AtBeginSection[]{%
	\begin{frame}<beamer>{Overview}
		\tableofcontents[currentsection]
	\end{frame}
}
\AtBeginSubsection[]{%
	\begin{frame}<beamer>{Overview}
		\tableofcontents[currentsection, currentsubsection]
	\end{frame}
}
\AtBeginSubsubsection[]{%
	\begin{frame}<beamer>{Overview}
    \begin{center}
      \insertsubsubsectionhead
    \end{center}
	\end{frame}
}
\begin{document}

@@

\begin{frame}{Tieto incident}
  \begin{itemize}
    \item In 2011 Tieto got a hardware problem in one of their datacenters. This
      created disruptions in many of the companies customers IT-system. For
      example: 
      \begin{itemize}
        \item All the pharmacies got problems handing out prescription medicin for a
          time.
        \item The swedish vehivle inspection agency could not report the results
          from the vehicle inspections they did during this time.
        \item A large logistical company could not handle its operation
        \item Social offices in Nacka and Sollentuna were unable to pay child
          support
      \end{itemize}
  \end{itemize}
\end{frame}
\begin{frame}{1177 incident}
  \begin{itemize}
    \item In 2019 a sub-contractor of a sub-contractor for a company that was
      handling the IT-system for three swedish regions health care guides (1177)
      call line, forgot to secure a backup server that stored the recordings
      from the phone calls.
    \item Resulted in around 2.7 million (around 170 000 hours) of recordings of
      confidential phone calls got accessible for the public.
    \item The recordings contains everything from medical symptoms to social
      security numbers.
  \end{itemize}
\end{frame}
\begin{frame}{The Swedish Transport Agency}
  \begin{itemize}
    \item In 2017 it was revealed that sensitive information from the Swedish
      Transport Agency database were made available for non-security screened
      employees abroad. 
    \item Information included:
      \begin{itemize}
        \item Driver licences (including picture)
        \item Company secrets
        \item Sensitive information about infrastructure (bridges, subways etc)
        \item Information that can be used to identify armoured cars.
      \end{itemize}
  \end{itemize}
\end{frame}
\begin{frame}{Swedish Public Employment Service}
  \begin{itemize}
    \item In 2019 it was found that the Swedish Publich Employment Services
      (arbetsförmedlingen) automatic system for reporting activity reports to
      the Swedish Social Insurance Agency (Försäkringskassan) sent incorrect
      reports.
    \item It was estimated that 10-15 percent of the decisions were incorrect.
    \item Resulted in that approximately 70 000 decisions could have been wrong.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Create commitment}
  \begin{itemize}
    \item The top management have the overall responsibility for the
      organisation -- This involves information security and security incidents.
    \item Important to make the management understand the importance of
      information security.
  \end{itemize}
\end{frame}

@@

\begin{frame}{ISO 27002 -- What should be done?}
  \begin{itemize}
    \item Security Policy.
    \item Organising the information security.
    \item Managing assets.
    \item Human resource and security.
    \item Physical and environmental security.
    \item Control communication and management.
    \item Access control.
    \item Acquisition, development and maintenance of information systems.
    \item Managing information security incidents.
    \item Continuity planning for the organisations.
    \item Compliance.
  \end{itemize}
\end{frame}

\section[Methodilogical support]{MSB:s methodological support}
\begin{frame}{Introduction}{\insertsectionhead}
  \begin{itemize}
    \item Swedish Civil Contingencies Agency (MSB).
    \item \enquote{MSB is responsible for issues concerning civil protection,
        public safety, emergency management and civil defence, as long as no
        other authority has responsibility\cite[About MSB]{msbse}\@.
        }
    \item This includes protection of our information systems.
  \end{itemize}
\end{frame}

@@

\begin{frame}{GDPR}{General Data Protection Regulation}
  \begin{itemize}
    \item Replacing Personal Data Act 1998:204
    \item Regulates how personal data are allowed to be stored and used.
      \begin{itemize}
        \item Protects a persons right of privacy\cite[art. 5]{eu:gdpr}
        \item Personal data can only be stored for as long as there is a need
          for it\cite[art. 5]{eu:gdpr}.
        \item Each person must give explicit consent to a company/organisation
          to store, publish or share personal data about them\cite[art.
          7]{eu:gdpr}.  
        \item Ensure that a person have the right to get access
          to all personal
          data that has been collected about them\cite[art. 15]{eu:gdpr}.
        \item Persons right to erasure\cite[art. 17]{eu:gdpr}. 
      \end{itemize}
  \end{itemize}
\end{frame}
\begin{frame}{GDPR cont.}{General Data Protection Regulation}
    \begin{itemize}
      \item Possibility to transfer personal data (data
          portability)\cite[art. 20]{eu:gdpr}.  
      \item Data protection by design and default \cite[art. 25]{eu:gdpr}.
      \item Records regarding processing activities must be maintained \cite[art.
        30]{eu:gdpr}.
      \item Personal data breaches must be reported to the supervisory
        authority, without undue delay, not later than 72 hours\cite[art.
        33]{eu:gdpr}.
    \end{itemize}
\end{frame}
\begin{frame}{GDPR cont.}{General Data Protection Regulation}
\begin{itemize}
    \item Individuals needs to be informed about the personal data
        breach\cite[art. 34]{eu:gdpr}.
    \item An independent agency should ensure that these rules are followed
      \cite[art 51]{eu:gdpr}.
    \item If the organisation does not comply to these regulations, they can be
      fined up to 4\% worldwide turnover or \euro{}20 Million, whichever is greater
      \cite[art 83]{eu:gdpr}.
    \item and much more.
\end{itemize}
\end{frame}

@@

    \midrule
    Considerable & 500 000 - 2 million SEK or 10-20\% over budget & Negative
    reporting in local and national media and in organised social media groups.
    Focused mainly on special instances or actions from a certain individual(s)
    & Disruption in one critical organisation that are longer than acceptable.
    \\ 
    \midrule
    Moderate & 1 - 500 000 or 5-10 \% over budget & A few dissatisfied
    individuals that express themselves in social medias or a smaller notice in
    local media.& Disruptions in one or several non-critical organisations that
    are longer than acceptable.\\ 
    \midrule
    Negligible & No loss & Little or non negative attention in media or social
    media & Negligable disruption in the organisation\\ 
    \bottomrule
  \end{tabularx}
  \end{table}
\end{frame}

>>>>

\DeclareAcronym{isms}{
  short = ISMS,
  long = Information Security Management Systems,
}

@@

\AtBeginSection[]{%
	\begin{frame}<beamer>{Overview}
		\tableofcontents[currentsection]
	\end{frame}
}
\AtBeginSubsection[]{%
	\begin{frame}<beamer>{Overview}
		\tableofcontents[currentsection, currentsubsection]
	\end{frame}
}
\AtBeginSubsubsection[]{%
	\begin{frame}<beamer>{Overview}
    \begin{center}
      \insertsubsubsectionhead{}
    \end{center}
	\end{frame}
}
\begin{document}

@@

\begin{frame}{Tieto incident}
  \begin{itemize}
    \item In 2011 Tieto got a hardware problem in one of their datacenters. This
      created disruptions in many of the companies customers IT-system. For
      example: 
      \begin{itemize}
        \item All the pharmacies got problems handing out prescription medicin for a
          time.
        \item The swedish vehivle inspection agency could not report the results
          from the vehicle inspections they did during this time.
        \item A large logistical company could not handle its operation
        \item Social offices in Nacka and Sollentuna were unable to pay child
          support
      \end{itemize}
  \end{itemize}
\end{frame}
\begin{frame}{1177 incident}
  \begin{itemize}
    \item In 2019 a sub-contractor of a sub-contractor for a company that was
      handling the IT-system for three swedish regions health care guides (1177)
      call line, forgot to secure a backup server that stored the recordings
      from the phone calls.
    \item Resulted in around 2.7 million (around 170 000 hours) of recordings of
      confidential phone calls got accessible for the public.
    \item The recordings contains everything from medical symptoms to social
      security numbers.
  \end{itemize}
\end{frame}
\begin{frame}{The Swedish Transport Agency}
  \begin{itemize}
    \item In 2017 it was revealed that sensitive information from the Swedish
      Transport Agency database were made available for non-security screened
      employees abroad. 
    \item Information included:
      \begin{itemize}
        \item Driver licences (including picture)
        \item Company secrets
        \item Sensitive information about infrastructure (bridges, subways etc)
        \item Information that can be used to identify armoured cars.
      \end{itemize}
  \end{itemize}
\end{frame}
\begin{frame}{Swedish Public Employment Service}
  \begin{itemize}
    \item In 2019 it was found that the Swedish Publich Employment Services
      (arbetsförmedlingen) automatic system for reporting activity reports to
      the Swedish Social Insurance Agency (Försäkringskassan) sent incorrect
      reports.
    \item It was estimated that 10 to 15 percent of the decisions were incorrect.
    \item Resulted in that approximately 70 000 decisions could have been wrong.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Create commitment}
  \begin{itemize}
    \item The top management have the overall responsibility for the
      organisation --- This involves information security and security incidents.
    \item Important to make the management understand the importance of
      information security.
  \end{itemize}
\end{frame}

@@

\begin{frame}{ISO 27002 --- What should be done?}
  \begin{itemize}
    \item Security Policy.
    \item Organising the information security.
    \item Managing assets.
    \item Human resource and security.
    \item Physical and environmental security.
    \item Control communication and management.
    \item Access control.
    \item Acquisition, development and maintenance of information systems.
    \item Managing information security incidents.
    \item Continuity planning for the organisations.
    \item Compliance.
  \end{itemize}
\end{frame}

\section[Methodilogical support]{MSB:s methodological support}
\begin{frame}{Introduction}{\insertsectionhead}
  \begin{itemize}
    \item Swedish Civil Contingencies Agency (MSB).
    \item \enquote{MSB is responsible for issues concerning civil protection,
        public safety, emergency management and civil defence, as long as no
        other authority has responsibility~\cite[About MSB]{msbse}\@.
        }
    \item This includes protection of our information systems.
  \end{itemize}
\end{frame}

@@

\begin{frame}{GDPR}{General Data Protection Regulation}
  \begin{itemize}
    \item Replacing Personal Data Act 1998:204
    \item Regulates how personal data are allowed to be stored and used.
      \begin{itemize}
        \item Protects a persons right of privacy~\cite[art. 5]{eu:gdpr}
        \item Personal data can only be stored for as long as there is a need
          for it~\cite[art. 5]{eu:gdpr}.
        \item Each person must give explicit consent to a company/organisation
          to store, publish or share personal data about them~\cite[art.
          7]{eu:gdpr}.  
        \item Ensure that a person have the right to get access
          to all personal
          data that has been collected about them~\cite[art. 15]{eu:gdpr}.
        \item Persons right to erasure~\cite[art. 17]{eu:gdpr}. 
      \end{itemize}
  \end{itemize}
\end{frame}
\begin{frame}{GDPR cont.}{General Data Protection Regulation}
    \begin{itemize}
      \item Possibility to transfer personal data (data
          portability)~\cite[art. 20]{eu:gdpr}.  
      \item Data protection by design and default~\cite[art. 25]{eu:gdpr}.
      \item Records regarding processing activities must be maintained~\cite[art.
        30]{eu:gdpr}.
      \item Personal data breaches must be reported to the supervisory
        authority, without undue delay, not later than 72 hours~\cite[art.
        33]{eu:gdpr}.
    \end{itemize}
\end{frame}
\begin{frame}{GDPR cont.}{General Data Protection Regulation}
\begin{itemize}
    \item Individuals needs to be informed about the personal data
        breach~\cite[art. 34]{eu:gdpr}.
    \item An independent agency should ensure that these rules are 
      followed~\cite[art 51]{eu:gdpr}.
    \item If the organisation does not comply to these regulations, they can be
      fined up to 4\% worldwide turnover or \euro{}20 Million, whichever is 
      greater~\cite[art 83]{eu:gdpr}.
    \item and much more.
\end{itemize}
\end{frame}

@@

    \midrule
    Considerable & 500 000 - 2 million SEK or 10-20\% over budget & Negative
    reporting in local and national media and in organised social media groups.
    Focused mainly on special instances or actions from a certain individual~(s)
    & Disruption in one critical organisation that are longer than acceptable.
    \\ 
    \midrule
    Moderate & 1 - 500 000 or 5-10 \% over budget & A few dissatisfied
    individuals that express themselves in social medias or a smaller notice in
    local media.& Disruptions in one or several non-critical organisations that
    are longer than acceptable.\\ 
    \midrule
    Negligible & No loss & Little or non negative attention in media or social
    media & Negligable disruption in the organisation\\ 
    \bottomrule
  \end{tabularx}
  \end{table}
\end{frame}
